CONEM2024-0025\\ NEW ABC MULTICRITERIA CLASSIFICATION WITH ANALYTICAL HIERARCHY PROCESS: HOW TO ATTRIBUTE WEIGHTS AND FORCE CONSISTENCY OF PAIRWISE COMPARISONS MATRIX

### 2024-02-29 accepted/CONEM2024/FCPCOM/CONEM2024.tex
updating

    % CABEÇALHO
    \thispagestyle{empty}
    \begin{figure}[h]
        \begin{center}
            \includegraphics[angle=0, width=\textwidth]{Logo_template.png}
        \end{center}
    \end{figure}
    \vspace{-.5cm}
    \hspace{-.8cm}
    \begin{tabular}{||p{\textwidth}}
    \begin{center}
    \vspace{-.6cm}
    %
    % Digitar o título
    \title{CONEM2024-0025\\ ABC  MULTICRITERIA CLASSIFICATION WITH ANALYTICAL HIERARCHY PROCESS: HOW TO FORCE CONSISTENCY OF PAIRWISE COMPARISONS MATRIX} 
    %
    \end{center}
	%
	% Digitar os nomes dos autores/instituição
    \authors{Tatiana Balbi Fraga, tatiana.balbi@ufpe.br$^1$} \\
    \authors{Beatriz Marinho Cavacanti, beatriz.marinhocavalcanti@ufpe.br$^1$} \\
    \authors{Alexia Maria Duque Silva, alexia.duque@ufpe.br$^1$} \\
    \authors{Erika Leticia Rodrigues Silva, erika.leticias@ufpe.br$^1$} \\\\
    \institution{$^1$Nome da instituição, endereço para correspondência (institucional, se houver)} \\
    \\
    %
    % Digitar o resumo
    \abstract{\textbf{Resumo:} Multicriteria classification is usually a crucial step in the decision-making in the manufacturing and purchasing management process. For such classification, the attribution of weights to the criteria, as well as the attribution of weights to the products according to each criterion, strongly influences the coherence of the categorization. Saaty's Analytic Hierarchy Process (AHP) is an eminent method for assigning weights to multiple criteria. AHP's logic is not complicated at all. However, since criteria's pairwise comparison matrices are usually generated manually and based only on some employee know-how, there is high complexity in elaborating a consistent pairwise matrix, especially when comparing several criteria. This paper presents a constructive algorithm that can adjust inconsistent matrices, forcing such matrices to have a better consistency rate and a new procedure and equation for ABC multicriteria classification. We tested the proposed solutions by applying the AHP method for ABC multicriteria classification to companies in three sectors. As a result, we observed that the algorithm for forcing consistency adjusts the pairwise matrices in just a few seconds, avoiding the manual work of several hours and that the procedures developed for multicriteria classification provided consistent results according to the analysis of company managers, showing to be essential resources for applying the AHP and ABC multicriteria classification methods.}\\
    %
    % Digitar as palavras-chave
   \\
    \keywords{\textbf{Palavras-chave:} ABC multicriteria classification, analytic hierarchy process, consistency rate,  constructive algorithm, COPSolver library}\\
    %
    \end{tabular}

@@

@@

  						& \multicolumn{7}{c | }{CR = 0.151}  & \multicolumn{7}{c}{CR = 0.096} \\
  						& \multicolumn{7}{c | }{$w=(0.05, 0.10, 0.14, 0.21, 0.02, 0.45, 0.02\underline{})$}  & \multicolumn{7}{c}{$w = (0.07, 0.16, 0.25, 0.20, 0.02, 0.26, 0.03)$} \\
   						& lt & sb & rp & cr & cm & bl & ob & lt & sb & rp & cr & cm & np & ob \\
						lt & 1.00 & 0.20 & 0.14 & \cellcolor[HTML]{ACE600} 0.20 & 7.00 & \cellcolor[HTML]{ACE600} 0.11 & 5.00 & 1.00 & 0.20 & 0.14 & \cellcolor[HTML]{ACE600} 0.33 & 7.00 & \cellcolor[HTML]{ACE600} 0.20 & 5.00 \\
						sb & 5.00 & 1.00 & 0.33 & \cellcolor[HTML]{ACE600} 0.33 & 5.00 &\cellcolor[HTML]{ACE600}  0.14 & 7.00 & 5.00 & 1.00 & 0.33 & \cellcolor[HTML]{ACE600} 1.00 & 5.00 & \cellcolor[HTML]{ACE600} 0.33 & 7.00 \\
						rp & 7.00 & 3.00 & 1.00 & \cellcolor[HTML]{ACE600} 0.33 & 5.00 & \cellcolor[HTML]{ACE600} 0.14 & 3.00 & 7.00 & 3.00 & 1.00 & \cellcolor[HTML]{ACE600} 1.00 & 7.00 & \cellcolor[HTML]{ACE600} 1.00 & 3.00 \\
						cr & \cellcolor[HTML]{ACE600} 5.00 & \cellcolor[HTML]{ACE600} 3.00 & \cellcolor[HTML]		{ACE600} 3.00 & 1.00 & \cellcolor[HTML]{ACE600} 9.00 & \cellcolor[HTML]{ACE600} 0.33 & 9.00 & \cellcolor[HTML]{ACE600} 3.00 & \cellcolor[HTML]{ACE600} 1.00 & \cellcolor[HTML]{ACE600} 1.00 & 1.00 & \cellcolor[HTML]{ACE600} 7.00 & \cellcolor[HTML]{ACE600} 1.00 & 9.00 \\
						cm & 0.14 & 0.20 & \cellcolor[HTML]{ACE600} 0.20 & \cellcolor[HTML]{ACE600} 0.11 & 1.00 & 0.11 & 1.00 & 0.14 & 0.20 & \cellcolor[HTML]{ACE600} 0.14 & \cellcolor[HTML]{ACE600} 0.14 & 1.00 & 0.11 & 1.00 \\
						np & \cellcolor[HTML]{ACE600} 9.00 & \cellcolor[HTML]{ACE600} 7.00 & \cellcolor[HTML]{ACE600} 7.00 & \cellcolor[HTML]{ACE600} 3.00 & 9.00 & 1.00 & 9.00 & \cellcolor[HTML]{ACE600} 5.00 & \cellcolor[HTML]{ACE600} 3.00 & \cellcolor[HTML]{ACE600} 1.00 & \cellcolor[HTML]{ACE600} 1.00 & 9.00 & 1.00 & 9.00 \\
						ob & 0.20 & 0.14 & 0.33 & 0.11 & 1.00 & 0.11 & 1.00 & 0.20 & 0.14 & 0.33 & 0.11 & 1.00 & 0.11 & 1.00 \\

@@

						\multicolumn{14}{l}{plastic packaging manufacturing company} \\

@@

   						& \multicolumn{7}{c | }{CR = 0.625}  & \multicolumn{7}{c}{CR = 0.085} \\
   						& \multicolumn{7}{c | }{$w=(0.39, 0.10, 0.22, 0.09, 0.12, 0.07)$}  & \multicolumn{7}{c}{$w = (0.50, 0.18, 0.14, 0.06, 0.09, 0.03)$} \\
   						& np & cr & lt & ob & sb & rp & & bl & cr & lt & ob & sb & rp & \\
						np & 1.00 & 3.00 & 7.00 & \cellcolor[HTML]{ACE600} 7.00 & 5.00 & \cellcolor[HTML]{ACE600} 5.00 & & 1.00 & 3.00 & 7.00 & \cellcolor[HTML]{ACE600} 9.00 & 5.00 & \cellcolor[HTML]{ACE600} 9.00 \\
						cr & 0.33 & 1.00 & 3.00 & \cellcolor[HTML]{ACE600} 0.20 & \cellcolor[HTML]{ACE600} 0.20 & \cellcolor[HTML]{ACE600} 0.20 & & 0.33 & 1.00 & 3.00 & \cellcolor[HTML]{ACE600} 3.00 & \cellcolor[HTML]{ACE600} 1.00 & \cellcolor[HTML]{ACE600} 5.00 \\
						lt & 0.14 & 0.33 & 1.00 & \cellcolor[HTML]{ACE600} 5.00 & \cellcolor[HTML]{ACE600} 7.00 & 7.00 & & 0.14 & 0.33 & 1.00 & \cellcolor[HTML]{ACE600} 3.00 & \cellcolor[HTML]{ACE600} 3.00 & 7.00 \\
						ob & \cellcolor[HTML]{ACE600} 0.14 & \cellcolor[HTML]{ACE600} 5.00 & \cellcolor[HTML]{ACE600} 0.20 & 1.00 & \cellcolor[HTML]{ACE600} 0.33 & 3.00 & & \cellcolor[HTML]{ACE600} 0.11 & \cellcolor[HTML]{ACE600} 0.33 & \cellcolor[HTML]{ACE600} 0.33 & 1.00 & \cellcolor[HTML]{ACE600} 1.00 & 3.00 \\
						sb & 0.20 & \cellcolor[HTML]{ACE600} 5.00 & \cellcolor[HTML]{ACE600} 0.14 & \cellcolor[HTML]{ACE600} 3.00 & 1.00 & 3.00 & & 0.20 & \cellcolor[HTML]{ACE600} 1.00 & \cellcolor[HTML]{ACE600} 0.33 & \cellcolor[HTML]{ACE600} 1.00 & 1.00 & 3.00 \\
						rp & \cellcolor[HTML]{ACE600} 0.20 & \cellcolor[HTML]{ACE600} 5.00 & 0.14 & 0.33 & 0.33 & 1.00 & & \cellcolor[HTML]{ACE600} 0.11 & \cellcolor[HTML]{ACE600} 0.20 & 0.14 & 0.33 & 0.33 & 1.00
				\end{tabular} \label{tab:pairwiseMatrix}
            \end{center}
	\end{table}

As we can see in this table, the pairwise comparison weights are adjusted primarily according to the weights assigned to the first three criteria. Thus, although there is a change in the pairwise comparison weights, which may be significant, the weights assigned to the first three criteria are preserved. The algorithm will change the weights of the other criteria to force the consistency of the pairwise comparisons matrix. Another significant observation is that the algorithm stops when it reaches the desired consistency rate. So, it is more likely that the algorithm changes the weights related to the fourth criterion and the next ones. Therefore, the ordering of criteria in the matrix must follow the following rules:

\begin{itemize}
\item the three most relevant criteria for the company are among the first;
\item the other criteria are ordered from the last to fourth position, in descending order of relevance.
\end{itemize}

For the data presented in Table 1, there was no concern with the correct ordering of the criteria. Based on the results and previous observations,  we requested the criteria classification according to its relevance to each company. Table 2 presents the modification of results after altering the criteria ordination on the input data.

@@

@@

    \section{REFERENCES}

    \section{INTRODUÇÃO}

        O objetivo destas instruções é servir de guia para a formatação dos trabalhos a serem publicados nos anais. Os anais do XII CONEM serão publicados no formato Adobe$^{\small{TM}}$ PDF.

        Os artigos \textcolor{red}{\underline{\textbf{DEVEM}}} ser formatados de acordo com estas instruções. O presente arquivo pode ser usado como modelo por usuários do \LaTeX. Além disso, pode ser usado como um guia de formatação para usuários de outros processadores de texto.

        Os trabalhos (\textit{draft} e \textit{final paper}) estão limitados a no mínimo \textcolor{red}{\underline{\textbf{6 páginas}}} e no máximo \textcolor{red}{\underline{\textbf{10 páginas}}}, incluindo tabelas e figuras. O arquivo final em formato PDF não deve exceder 2 MB.

        A língua oficial do congresso é o português; entretanto serão aceitos manuscritos em espanhol ou em inglês. Se o trabalho não for escrito em inglês, o autor deverá incluir o título, os nomes dos autores e afiliações, o resumo e as palavras-chave, traduzidos para o inglês, após a lista de referências, no fim do artigo.

    \section{FORMATO DO TEXTO}

        O artigo deve ser digitado no \textit{template} ``CONEM2024.tex'' e as principais configurações (ex.: formato do papel, tipo e tamanho da fonte do título/Seção/Subseção, espaços, margens, alinhamentos, indentação, etc.) já estão previamente configuradas no arquivo ``abcm.sty''. Título, nomes dos autores, instituição, endereço, resumo e palavras-chave devem ser recuados 0,1 cm da margem esquerda e marcados por uma linha preta na borda esquerda com largura de 2$\frac{1}{4}$ pontos.

        Informações suficientes devem ser fornecidas diretamente no texto, ou por referência a trabalhos publicados disponíveis. Notas de rodapé devem ser evitadas.

>>>>

    % CABEÇALHO
    \thispagestyle{empty}
    \begin{figure}[h]
        \begin{center}
            \includegraphics[angle=0, width=\textwidth]{Logo_template.png}
        \end{center}
    \end{figure}
    \vspace{-.5cm}
    \hspace{-.8cm}
    \begin{tabular}{||p{\textwidth}}
    \begin{center}
    \vspace{-.6cm}
    %
    % Digitar o título
    \title{CONEM2024-0025\\ NEW ABC  MULTICRITERIA CLASSIFICATION WITH ANALYTICAL HIERARCHY PROCESS: HOW TO ATTRIBUTE WEIGHTS AND FORCE CONSISTENCY OF PAIRWISE COMPARISONS MATRIX} 
    %
    \end{center}
	%
	% Digitar os nomes dos autores/instituição
    \authors{Tatiana Balbi Fraga, tatiana.balbi@ufpe.br$^1$} \\
    \authors{Beatriz Marinho Cavacanti, beatriz.marinhocavalcanti@ufpe.br$^1$} \\
    \authors{Alexia Maria Duque Silva, alexia.duque@ufpe.br$^1$} \\
    \authors{Erika Leticia Rodrigues Silva, erika.leticias@ufpe.br$^1$} \\\\
    \institution{$^1$Nome da instituição, endereço para correspondência (institucional, se houver)} \\
    \\
    %
    % Digitar o resumo
    \abstract{\textbf{Resumo:} Multicriteria classification is usually a crucial step in the decision-making in the manufacturing and purchasing management process. For such classification, the attribution of weights to the criteria, as well as the attribution of weights to the products according to each criterion, strongly influences the coherence of the categorization. Saaty's Analytic Hierarchy Process (AHP) is an eminent method for assigning weights to multiple criteria. AHP's logic is not complicated at all. However, since criteria's pairwise comparison matrices are usually generated manually and based only on some employee know-how, there is high complexity in elaborating a consistent pairwise matrix, especially when comparing several criteria. This paper presents a constructive algorithm that can adjust inconsistent matrices, forcing such matrices to have a better consistency rate and a new procedure and equation for ABC multicriteria classification. We tested the proposed solutions by applying the AHP method for ABC multicriteria classification to companies in three sectors. As a result, we observed that the algorithm for forcing consistency adjusts the pairwise matrices in just a few seconds, avoiding the manual work of several hours and that the new solution method developed for multicriteria classification provided consistent results according to the analysis of company managers, showing to be essential resources for applying the AHP and ABC multicriteria classification methods.}\\
    %
    % Digitar as palavras-chave
   \\
    \keywords{\textbf{Palavras-chave:} ABC multicriteria classification, analytic hierarchy process, consistency rate,  constructive algorithm, COPSolver library}\\
    %
    \end{tabular}

@@

                    	\multicolumn{7}{c}{original data} & & \multicolumn{7}{c}{adjusted} \\

@@

  						& \multicolumn{7}{c | }{CR = 0.136}  & \multicolumn{7}{c}{CR = 0.096} \\
  						& \multicolumn{7}{c | }{$w=(0.07, 0.13, 0.26, 0.03, 0.48, 0.03)$}  & \multicolumn{7}{c}{$w = (0.09, 0.16, 0.35, 0.03, 0.35, 0.03)$} \\
   						& lt & rp & cr & cm & bl & ob & & lt & rp & cr & cm & bl & ob \\
						lt & 1.00 & 0.20 & 0.14 & 7.00 & \cellcolor[HTML]{ACE600} 0.11 & 5.00 & & 1.00 & 0.20 & 0.14 & 7.00 & \cellcolor[HTML]{ACE600} 0.14 & 5.00 \\
						rp & 5.00 & 1.00 & 0.33 & 5.00 & \cellcolor[HTML]{ACE600} 0.14 & 3.00 & & 5.00 & 1.00 & 0.33 & 5.00 & \cellcolor[HTML]{ACE600} 0.33 & 3.00 \\
						cr & 7.00 & 3.00 & 1.00 & 9.00 & \cellcolor[HTML]{ACE600} 0.33 & 9.00 & & 7.00 & 3.00 & 1.00 & 9.00 & \cellcolor[HTML]{ACE600} 1.00 & 9.00 \\
						cm & 0.14 & 0.20 & 0.11 & 1.00 & 0.11 & 1.00 & & 0.14 & 0.20 & 0.11 & 1.00 & 0.11 & 1.00 \\
						bl & \cellcolor[HTML]{ACE600} 9.00 & \cellcolor[HTML]{ACE600} 7.00 & \cellcolor[HTML]{ACE600} 3.00 & 9.00 & 1.00 & 9.00 & & \cellcolor[HTML]{ACE600} 7.00 & \cellcolor[HTML]{ACE600} 3.00 & \cellcolor[HTML]{ACE600} 1.00 & 9.00 & 1.00 & 9.00 \\
						ob & 0.20 & 0.33 & 0.11 & 1.00 & 0.11 & 1.00 & & 0.20 & 0.33 & 0.11 & 1.00 & 0.11 & 1.00 \\

@@

						\multicolumn{14}{l}{plastic packaging manufacturing company} \\
						\cline {1-15} \\
   						& \multicolumn{7}{c | }{CR = 0.625}  & \multicolumn{7}{c}{CR = 0.085} \\
   						& \multicolumn{7}{c | }{$w=(0.39, 0.10, 0.22, 0.09, 0.12, 0.07)$}  & \multicolumn{7}{c}{$w = (0.50, 0.18, 0.14, 0.06, 0.09, 0.03)$} \\
   						& np & cr & lt & ob & sb & rp & & bl & cr & lt & ob & sb & rp & \\
						np & 1.00 & 3.00 & 7.00 & \cellcolor[HTML]{ACE600} 7.00 & 5.00 & \cellcolor[HTML]{ACE600} 5.00 & & 1.00 & 3.00 & 7.00 & \cellcolor[HTML]{ACE600} 9.00 & 5.00 & \cellcolor[HTML]{ACE600} 9.00 \\
						cr & 0.33 & 1.00 & 3.00 & \cellcolor[HTML]{ACE600} 0.20 & \cellcolor[HTML]{ACE600} 0.20 & \cellcolor[HTML]{ACE600} 0.20 & & 0.33 & 1.00 & 3.00 & \cellcolor[HTML]{ACE600} 3.00 & \cellcolor[HTML]{ACE600} 1.00 & \cellcolor[HTML]{ACE600} 5.00 \\
						lt & 0.14 & 0.33 & 1.00 & \cellcolor[HTML]{ACE600} 5.00 & \cellcolor[HTML]{ACE600} 7.00 & 7.00 & & 0.14 & 0.33 & 1.00 & \cellcolor[HTML]{ACE600} 3.00 & \cellcolor[HTML]{ACE600} 3.00 & 7.00 \\
						ob & \cellcolor[HTML]{ACE600} 0.14 & \cellcolor[HTML]{ACE600} 5.00 & \cellcolor[HTML]{ACE600} 0.20 & 1.00 & \cellcolor[HTML]{ACE600} 0.33 & 3.00 & & \cellcolor[HTML]{ACE600} 0.11 & \cellcolor[HTML]{ACE600} 0.33 & \cellcolor[HTML]{ACE600} 0.33 & 1.00 & \cellcolor[HTML]{ACE600} 1.00 & 3.00 \\
						sb & 0.20 & \cellcolor[HTML]{ACE600} 5.00 & \cellcolor[HTML]{ACE600} 0.14 & \cellcolor[HTML]{ACE600} 3.00 & 1.00 & 3.00 & & 0.20 & \cellcolor[HTML]{ACE600} 1.00 & \cellcolor[HTML]{ACE600} 0.33 & \cellcolor[HTML]{ACE600} 1.00 & 1.00 & 3.00 \\
						rp & \cellcolor[HTML]{ACE600} 0.20 & \cellcolor[HTML]{ACE600} 5.00 & 0.14 & 0.33 & 0.33 & 1.00 & & \cellcolor[HTML]{ACE600} 0.11 & \cellcolor[HTML]{ACE600} 0.20 & 0.14 & 0.33 & 0.33 & 1.00
				\end{tabular} \label{tab:pairwiseMatrix}
            \end{center}
	\end{table}

As we can see in this table, the pairwise comparison weights are adjusted primarily according to the weights assigned to the first three criteria. Thus, although there is a change in the pairwise comparison weights, which may be significant, the weights assigned to the first three criteria are preserved. The algorithm will change the weights of the other criteria to force the consistency of the pairwise comparisons matrix. Another significant observation is that the algorithm stops when it reaches the desired consistency rate. So, it is more likely that the algorithm changes the weights related to the fourth criterion and the next ones. Therefore, in the preparation of the input file the data of the criteria must be informed in such a way that the criteria are ordered in descending order according to their respective relevance.

For the data presented in Table \ref{tab:pairwiseMatrix}, there was no concern with the correct ordering of the criteria. Based on the results and previous observations, we have reordered the criteria using their respective weights, initially defined by the normalized eigenvector of the original pairwise comparisons matrix ($w$ in Tab. \ref{tab:pairwiseMatrix}). Table \ref{tab:pairwiseMatrixReordered} presents the modification of results after altering the criteria ordination on the input data.
   						\multicolumn{7}{c}{original data reordered} & & \multicolumn{7}{c}{adjusted} \\

@@

  						& \multicolumn{7}{c | }{CR = 0.136}  & \multicolumn{7}{c}{CR = 0.082} \\
  						& \multicolumn{7}{c | }{$w=(0.48, 0.26, 0.13, 0.07, 0.03, 0.03)$}  & \multicolumn{7}{c}{$w = (0.49, 0.27, 0.11, 0.06, 0.05, 0.03)$} \\
   						& bl & cr & rp & lt & ob & cm & & bl & cr & rp & lt & ob & cm \\        
						bl & 1.00 & 3.00 & 7.00 & 9.00 & 9.00 & 9.00 & & 1.00 & 3.00 & 7.00 & 9.00 & 9.00 & 9.00 \\
						cr & 0.33 & 1.00 & 3.00 & 7.00 & 9.00 & 9.00 & & 0.33 & 1.00 & 3.00 & 7.00 & 9.00 & 9.00 \\
						rp & 0.14 & 0.33 & 1.00 & \cellcolor[HTML]{ACE600} 5.00 & 3.00 & 5.00 & & 0.14 & 0.33 & 1.00 & \cellcolor[HTML]{ACE600} 3.00 & 3.00 & 5.00 \\
						lt & 0.11 & 0.14 & \cellcolor[HTML]{ACE600} 0.20 & 1.00 & \cellcolor[HTML]{ACE600} 5.00 & 7.00 & & 0.11 & 0.14 & \cellcolor[HTML]{ACE600} 0.33 & 1.00 & \cellcolor[HTML]{ACE600} 1.00 & 7.00 \\
						ob & 0.11 & 0.11 & 0.33 & \cellcolor[HTML]{ACE600} 0.20 & 1.00 & \cellcolor[HTML]{ACE600} 1.00 & & 0.11 & 0.11 & 0.33 & \cellcolor[HTML]{ACE600} 1.00 & 1.00 & \cellcolor[HTML]{ACE600} 3.00 \\
						sb & 0.11 & 0.11 & 0.20 & 0.14 & \cellcolor[HTML]{ACE600} 1.00 & 1.00 & & 0.11 & 0.11 & 0.20 & 0.14 & \cellcolor[HTML]{ACE600} 0.33 & 1.00 \\

						\multicolumn{14}{l}{plastic packaging manufacturing company} \\
						\cline {1-15} \\
   						& \multicolumn{7}{c | }{CR = 0.625}  & \multicolumn{7}{c}{CR = 0.094} \\
   						& \multicolumn{7}{c | }{$w=(0.40, 0.22, 0.12, 0.07, 0.10, 0.10)$}  & \multicolumn{7}{c}{$w = (0.56, 0.21, 0.08, 0.07, 0.07, 0.02)$} \\
   						& bl & lt & sb & rp & ob & cr & & bl & lt & sb & rp & ob & cr & \\
						bl & 1.00 & 7.00 & 5.00 & \cellcolor[HTML]{ACE600} 5.00 & 7.00 & \cellcolor[HTML]{ACE600} 3.00 & & 1.00 & 7.00 & 5.00 & \cellcolor[HTML]{ACE600} 9.00 & 7.00 & \cellcolor[HTML]{ACE600} 9.00 \\
						lt & 0.14 & 1.00 & \cellcolor[HTML]{ACE600} 7.00 & \cellcolor[HTML]{ACE600} 7.00 & 5.00 & \cellcolor[HTML]{ACE600} 0.33 & & 0.14 & 1.00 & \cellcolor[HTML]{ACE600} 3.00 & \cellcolor[HTML]{ACE600} 5.00 & 5.00 & \cellcolor[HTML]{ACE600} 9.00 \\
						sb & 0.20 & \cellcolor[HTML]{ACE600} 0.14 & 1.00 & \cellcolor[HTML]{ACE600} 3.00 & \cellcolor[HTML]{ACE600} 3.00 & 5.00 & & 0.20 & \cellcolor[HTML]{ACE600} 0.33 & 1.00 & \cellcolor[HTML]{ACE600} 1.00 & \cellcolor[HTML]{ACE600} 1.00 & 5.00 \\
						rp & \cellcolor[HTML]{ACE600} 0.20 & \cellcolor[HTML]{ACE600} 0.14 & \cellcolor[HTML]{ACE600} 0.33 & 1.00 & \cellcolor[HTML]{ACE600} 0.33 & 5.00 & & \cellcolor[HTML]{ACE600} 0.11 & \cellcolor[HTML]{ACE600} 0.20 & \cellcolor[HTML]{ACE600} 1.00 & 1.00 & \cellcolor[HTML]{ACE600} 1.00 & 5.00 \\
						ob & 0.14 & 0.20 & \cellcolor[HTML]{ACE600} 0.33 & \cellcolor[HTML]{ACE600} 3.00 & 1.00 & 5.00 & & 0.14 & 0.20 & \cellcolor[HTML]{ACE600} 1.00 & \cellcolor[HTML]{ACE600} 1.00 & 1.00 & 5.00 \\
						cr & \cellcolor[HTML]{ACE600} 0.33 & \cellcolor[HTML]{ACE600} 3.00 & 0.20 & 0.20 & 0.20 & 1.00 & & \cellcolor[HTML]{ACE600} 0.11 & \cellcolor[HTML]{ACE600} 0.11 & 0.20 & 0.20 & 0.20 & 1.00 \\
				\end{tabular} \label{tab:pairwiseMatrixReordered}
            \end{center}
\end{table}

a Tab. \ref{tab:pairwiseMatrixAdjusted} apresenta uma comparação dos resultados usando diferentes métodos. Na matrix PCM 02, o método para forçar consistência foi aplicado antes da reordenação dos critérios antes. Já no caso da matrix PCM 03, o método para forçar consistência foi aplicado após a reordenação dos critérios. Esta tabela também mostra a matrix PCM 04 com os ajustes realizados pelo dono da empresa na matrix PCM 01 após verificar os resultados encontrados pelo COPSolver. Conforme podemos verificar, o software COPSolver provê uma importante ajuda na elaboração de tabelas de comparação por pares consistentes. Verificamos também que é importante realizar a ordenação dos critérios antes de ajustar a matriz buscando uma melhor consistência, pois os pesos atribuídos para os critérios mais relevantes devem ser preferencialmente mantidos. Finalmente, observamos que a ordenação dos critérios também ajuda no ajuste manual das matrizes de comparação por pares, já que a apresentação destas tabelas com os critérios ordenados facilita a análise dos pesos.  Para classificação multicritério dos itens foi utilizada a matriz de comparação por pares PCM 04, tendo em vista a importância da avaliação final realizada pelo funcionário responsável ou dono da empresa.

\begin{table}[ht]
            \begin{center}
                \caption{\textbf{Pairwise comparisons matrix and consistency rate changes for three companies 					(results found by COPSolver) - legend: sb = substitutability; lt = lead-time; rp = repairability; cr = criticality; ob = obsolescence; bl = billing; cm = commonality; CR = consistence rate; w = normalized weights vector.}}
                    \begin{tabular}[l]{p{0.4cm} p{0.5cm} p{0.5cm} p{0.5cm} p{0.5cm} p{0.5cm} p{0.5cm} p{0.5cm} | p{0.5cm} p{0.5cm} p{0.5cm} p{0.5cm} p{0.5cm} p{0.5cm} p{0.5cm}}
                        \\
   						\multicolumn{14}{c}{furniture trades company} \\
   						\cline {1-15} \\
   						\multicolumn{7}{l}{PCM 01 - original data reordered} & & \multicolumn{7}{l}{PCM 02 - PCM 01 adjusted and than reordered} \\
						\cline {1-15} \\
  						& \multicolumn{7}{c | }{CR = 0.136}  & \multicolumn{7}{c}{CR = 0.096} \\
  						& \multicolumn{7}{c | }{$w=(0.48, 0.26, 0.13, 0.07, 0.03, 0.03)$}  & \multicolumn{7}{c}{$w = (0.35, 0.35, 0.16, 0.09, 0.03, 0.03)$} \\
   						& bl & cr & rp & lt & ob & cm & & bl & cr & rp & lt & ob & cm \\        
						bl & 1.00 & 3.00 & 7.00 & 9.00 & 9.00 & 9.00 & & 1.00 & \cellcolor[HTML]{ACE600} 1.00 & \cellcolor[HTML]{ACE600} 3.00 & \cellcolor[HTML]{ACE600} 7.00 & 9.00 & 9.00 \\
						cr & 0.33 & 1.00 & 3.00 & 7.00 & 9.00 & 9.00 & & \cellcolor[HTML]{ACE600} 1.00 & 1.00 & 3.00 & 7.00 & 9.00 & 9.00 \\
						rp & 0.14 & 0.33 & 1.00 & 5.00 & 3.00 & 5.00 & & \cellcolor[HTML]{ACE600} 0.33 & 0.33 & 1.00 & 5.00 & 3.00 & 5.00 \\
						lt & 0.11 & 0.14 & 0.20 & 1.00 & 5.00 & 7.00 & & \cellcolor[HTML]{ACE600} 0.14 & 0.14 & 0.20 & 1.00 & 5.00 & 7.00 \\
						ob & 0.11 & 0.11 & 0.33 & 0.20 & 1.00 & 1.00 & & 0.11 & 0.11 & 0.33 & 0.20 & 1.00 & 1.00 \\
						sb & 0.11 & 0.11 & 0.20 & 0.14 & 1.00 & 1.00 & & 0.11 & 0.11 & 0.20 & 0.14 & 1.00 & 1.00 \\
						\cline {1-15} \\
						\multicolumn{7}{l}{PCM 03 - PCM 01 reordered and than adjusted} & & \multicolumn{7}{l}{PCM 01 - manually adjusted based on COPSolver results} \\
						\cline {1-15} \\
  						& \multicolumn{7}{c | }{CR = 0.082}  & \multicolumn{7}{c}{CR = 0.082} \\
  						& \multicolumn{7}{c | }{$w=(0.49, 0.27, 0.11, 0.06, 0.05, 0.03)$}  & \multicolumn{7}{c}{$w = (0.50, 0.25, 0.12, 0.08, 0.03, 0.03)$} \\
   						& bl & cr & rp & lt & ob & cm & & bl & cr & rp & lt & ob & cm \\        
						bl & 1.00 & 3.00 & 7.00 & 9.00 & 9.00 & 9.00 & & 1.00 & 3.00 & 7.00 & 9.00 & 9.00 & 9.00 \\
						cr & 0.33 & 1.00 & 3.00 & 7.00 & 9.00 & 9.00 & & 0.33 & 1.00 & 3.00 & \cellcolor[HTML]{ACE600} 5.00 & 9.00 & 9.00 \\
						rp & 0.14 & 0.33 & 1.00 & \cellcolor[HTML]{ACE600} 3.00 & 3.00 & 5.00 & & 0.14 & 0.33 & 1.00 & \cellcolor[HTML]{ACE600} 3.00 & \cellcolor[HTML]{ACE600} 5.00 & 5.00 \\
						lt & 0.11 & 0.14 & \cellcolor[HTML]{ACE600} 0.33 & 1.00 & \cellcolor[HTML]{ACE600} 1.00 & 7.00 & & 0.11 & \cellcolor[HTML]{ACE600} 0.20 & \cellcolor[HTML]{ACE600} 0.33 & 1.00 & 5.00 & \cellcolor[HTML]{ACE600} 5.00 \\
						ob & 0.11 & 0.11 & 0.33 & \cellcolor[HTML]{ACE600} 1.00 & 1.00 & \cellcolor[HTML]{ACE600} 3.00 & & 0.11 & 0.11 & \cellcolor[HTML]{ACE600}  0.20 & 0.20 & 1.00 & 1.00 \\
						sb & 0.11 & 0.11 & 0.20 & 0.14 & \cellcolor[HTML]{ACE600} 0.33 & 1.00 & & 0.11 & 0.11 & 0.20 & \cellcolor[HTML]{ACE600} 0.20 & 1.00 & 1.00 \\
						\cline {1-15} \\
				\end{tabular} \label{tab:pairwiseMatrixAdjusted}
            \end{center}
	\end{table}

@@

\subsection{Analysis on the New ABC Multicriteria Classification Method}

\begin{table}[h!]
\begin{center}
\begin{small}
\begin{tabular}[c]{l r r | l r r r r}
\multicolumn{8}{l}{car mechanics company} \\
\cline {1-8} \\
\multicolumn{3}{l |}{multc. ABC clssf.} & \multicolumn{5}{c}{demand pattern classification}\\
\multicolumn{3}{c |}{cutoff}  & \multicolumn{5}{c}{75 sales between October 2020 and July 2023} \\
\multicolumn{3}{c |}{0.35 \%; 0.70 \%} & & \multicolumn{1}{c}{total} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} & \multicolumn{1}{c}{C}\\
A &  10 & 23.81 \% & smooth & 18.75 \% &  20.00 \% & 16.67 \% & 0 \%\\
B &  15 & 35.71 \% & slow-moving & 6.25 \% & 10.00 \% & 0.00 \% & 0 \%\\
C &  17 & 40.48 \% & sporadic & 75.00 \% & 70.00 \% & 83.33 \% & 0 \% \\
\cline {1-8} \\
\multicolumn{8}{l}{furniture trades company} \\
\cline {1-8} \\
\multicolumn{3}{l |}{multc. ABC clssf.} & \multicolumn{5}{c}{demand pattern classification}\\
\multicolumn{3}{c |}{cutoff:} &\\
\multicolumn{3}{c |}{0.35 \%; 0.70 \%}& & \multicolumn{1}{c}{total} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} & \multicolumn{1}{c}{C}\\
A &  26 & 10.24 \% & smooth & 31.58 \% &  73.08 \% & 30.00 \% & 24.82 \%\\
B &  80 & 31.50 \% & slow-moving & 18.62 \% & 19.23 \% & 33.75 \% & 9.93 \%\\
C & 148 & 58.27 \% & sporadic & 49.80 \% & 7.69 \% & 36.25 \% & 65.25 \% \\
\cline {1-8} \\
 \multicolumn{8}{l}{plastic packaging manufacturing company}\\
\cline {1-8} \\
\multicolumn{3}{l |}{multc. ABC clssf.} & \multicolumn{5}{c}{demand pattern classification}\\
\multicolumn{3}{c |}{cutoff:} & \multicolumn{5}{c}{5,967 sales between November 2019 and February 2023}\\
\multicolumn{3}{c |}{0.35 \%; 0.70 \%} & & \multicolumn{1}{c}{total} & \multicolumn{1}{c}{A} & \multicolumn{1}{c}{B} & \multicolumn{1}{c}{C}\\
A &   128 &  5.49 \% & smooth & 15.31 \% & 22.66 \% & 23.97 \% & 10.42 \%\\
B &   726 & 31.12 \% & slow-moving & 5.53 \% & 1.56 \% & 8.68 \% & 4.33 \%\\
C & 1,479 & 63.39 \% & sporadic & 79.16 \% & 75.78 \% & 67.36 \% & 85.25 \% \\
\cline {1-8} \\

\end{tabular}
\caption{ABC multicriteria and demand pattern classifications for three companies (results found by COPSolver)}
\label{tab:results}
\end{small}
\end{center}
\end{table}

@@
